Statistical Searches for Microlensing Events in Large, Non-Uniformly Sampled Time-Domain Surveys: A Test Using Palomar Transient Factory Data

### 2013-10-01 text/final_paper/ms.tex
removed reference to in prep work
removed footnote to bitly

In this paper, we use the Palomar Transient Factory (PTF) survey, described in Section~\ref{sec:ptf}, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys. Although to date no microlensing event has been reported using PTF data, a simple extrapolation from previous all-sky rate estimates suggests $\sim$1--10 detectable microlensing events within the PTF footprint over the survey's three years of operation \citep{han2008, gaudi2008}. We examine the selection methods used by microlensing surveys to identify event signatures in light curves as well as an event identification procedure based on a set of variability statistics. We then present the detection efficiency for these methods as computed for a representative set of PTF fields (Section~\ref{sec:event_recovery}). In Section~\ref{sec:search}, we apply a set of criteria that use the results of these tests to all PTF light curves with $>$10 high-quality observations to identify candidate microlensing events. We discuss eight interesting transients and three candidate events identified in this manner before concluding in Section~\ref{sec:conc}. A detailed investigation of the all-sky microlensing event rate,  generated from a model that includes realistic mass and velocity distributions, will be presented in Fournier et al. (in prep).

@@

As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint}). The PTF survey footprint is not uniformly sampled either spatially or temporally.\footnote{For an interactive visualization of the PTF survey footprint, see \url{http://bit.ly/PTFSky}.} Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler. The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}. The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry. Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.

>>>>

In this paper, we use the Palomar Transient Factory (PTF) survey, described in Section~\ref{sec:ptf}, as a test case for studying the frequency and detectability of microlensing events in all-sky, irregularly sampled, time-domain surveys. Although to date no microlensing event has been reported using PTF data, a simple extrapolation from previous all-sky rate estimates suggests $\sim$1--10 detectable microlensing events within the PTF footprint over the survey's three years of operation \citep{han2008, gaudi2008}. We examine the selection methods used by microlensing surveys to identify event signatures in light curves as well as an event identification procedure based on a set of variability statistics. We then present the detection efficiency for these methods as computed for a representative set of PTF fields (Section~\ref{sec:event_recovery}). In Section~\ref{sec:search}, we apply a set of criteria that use the results of these tests to all PTF light curves with $>$10 high-quality observations to identify candidate microlensing events. We discuss eight interesting transients and three candidate events identified in this manner before concluding in Section~\ref{sec:conc}. % A detailed investigation of the all-sky microlensing event rate,  generated from a model that includes realistic mass and velocity distributions, will be presented in Fournier et al. (in prep).

@@

As of 2012 Dec, the PTF footprint includes $\sim$15,224 (2766) deg$^2$ imaged $>$10 ($>$100) times in $R$-band and $\sim$5430 (290) deg$^2$ imaged that often in $g$-band (see Figure~\ref{fig:survey_footprint}). The PTF survey footprint is not uniformly sampled either spatially or temporally. Each field has a unique sampling pattern determined by which of the PTF sub-surveys it belongs to, what time of the year it is visible, and how high a priority it is given by the scheduler. The imaging cadences range from 1-5 days, but there is a higher-cadence campaign of a field in Orion to find transiting planets around young stars \citep{nick2009}. The result is that PTF light curves often contain gaps and regions of high-cadence observations and/or of low-cadence observations, so that the archive is a massive dataset of irregularly-sampled, time-domain photometry. Figure~\ref{fig:sampling} shows six randomly selected light curves and illustrates the varying cadences and coverage that different fields may have over the same one-year period.
